Structural analysis

### 2024-12-06 workflow/notebooks/draft.Rmd
update README

p_dbs <- ggplot(jaccard_df, aes(avg_Jaccard, 1, color=singleton, alpha=db)) +
  stat_ecdf(pad=FALSE) +
  scale_color_manual(values=palette_singleton_p) +
  # scale_linetype_manual(values = c(2,2,2,2,1)) +
  scale_alpha_manual(values = seq(0.3, 1, length.out=4)) +
  coord_fixed() +
  scale_y_continuous(expand = expansion(0.0,0)) +  
  scale_x_continuous(expand = expansion(add = c(0,0.01))) +  
  guides(colour = guide_legend(order = 1),  
         alpha = guide_legend(order = 2)) +
  theme(panel.border = element_rect(fill = NA), 
        legend.position = "bottom",
        panel.grid.major = element_line(linetype = 2)) + 
  # theme_bw(base_family = "Aptos") +
  labs(x="Mean Jaccard Index", y="ECDF", 
       alpha="DB", linetype="DB", color="Target sets")

@@

```{r fig3}
xax <- ifelse(df_model$algorithm=="Dist.", "italic", "plain")

plot_DL <- reco %>% 
  left_join(df_model) %>% 
  ggplot(aes(weave_factors(model, data), n_events, 
             color=algorithm, fill=targets)) + 
  labs(y="(D+L)/# Tips", x="",fill="Target sets") +
  geom_boxplot(outlier.size = .1) +
  scale_color_manual(values = c("black", "black"),
                     labels = c("ML", expression(italic("Dist.")))) +
  guides(x = "axis_nested", 
         colour=guide_legend(title = "Method", override.aes=list(linetype = 0, size=0))) +
  scale_fill_manual(values = palettes_method) +
    theme(
    axis.text.x = element_text(face = xax),
    ggh4x.axis.nestline.x = element_line(color="grey"),
    ggh4x.axis.nesttext.x = element_text(colour = "grey30", 
                                         face = "bold", size=10)
  )

plot_rf <- disco_rf %>% 
  mutate(model=factor(model, levels=models)) %>% 
  filter(nsps>length(sptree$tip.label)/2) %>%
  # separate(bn, c("seed", "targets", "alphabet", "model")) %>% 
  left_join(df_model) %>% 
  ggplot(aes(weave_factors(model, data), rf, 
             color=algorithm, fill=target)) + 
  labs(x="", y="Norm. RF to species tree", fill="Target sets") +
  geom_boxplot(outlier.size = .1) +
  scale_color_manual(values = c("black", "black"),
                     labels = c("ML", expression(italic("Dist.")))) +
  guides(x = "axis_nested", 
         colour=guide_legend(title = "Method", override.aes=list(linetype = 0, size=0))) +
  scale_fill_manual(values = palettes_method) +
    theme(
    axis.text.x = element_text(face = xax),
    ggh4x.axis.nestline.x = element_line(color="grey"),
    ggh4x.axis.nesttext.x = element_text(colour = "grey30", 
                                         face = "bold", size=10)
  )

@@

plot_apro <- trees_df %>% 
  left_join(df_model) %>% 
  ggplot(aes(weave_factors(model, data), q1, 
             color=algorithm,fill=targets)) + 
  labs(x="", y="Quartet support", fill="Target sets") + 
  geom_boxplot(outlier.size = .1) +
  scale_color_manual(values = c("black", "black"),
                     labels = c("ML", expression(italic("Dist.")))) +
  guides(x = "axis_nested", 
         colour=guide_legend(title = "Method", override.aes=list(linetype = 0, size=0))) +
  scale_fill_manual(values = palettes_method) +
    theme(
    axis.text.x = element_text(face = xax),
    ggh4x.axis.nestline.x = element_line(color="grey"),
    ggh4x.axis.nesttext.x = element_text(colour = "grey30", 
                                         face = "bold", size=10)
  )

@@

```{r FigS1}
# Compare trees with RF distance
rf_df <- get_rf_df(df_trees)
rf_text <- group_by(rf_df, model.x, model.y) %>% 
  # filter(as.integer(model.x)<=as.integer(model.y)) %>% 
  summarise(value=mean(RF, na.rm=T))

@@

rf_plot <- rf_text %>% 
  ggplot(aes(model.x, model.y)) +
  geom_tile(aes( fill=value)) +
  geom_text(aes(label=round(value, 2)), color="white") +
  scale_fill_distiller(palette="YlGnBu", direction = -1, limits = c(0,1), 
                       name="Mean RF") + 
  labs(x="Mean RF Distance", y="") +
  theme_classic() +
  theme(legend.position = "none", 
        axis.line = element_blank(),
        axis.ticks = element_blank())

@@

fig_s1 <- ((plot_bs / rf_plot) + 
             plot_layout(heights = c(2,1)) | 
             (plot_dist / var_r2t) + 
             plot_layout(heights = c(2,1))) +
  plot_layout(widths = c(1, 2)) +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))

ggsave("draft/figures/Fig_S1.pdf", fig_s1, width = 10, height = 6)
```

```{r FigS2}
# pp1_plot <- nodes_df %>%
#     filter(model!="FTPY") %>% 
#     ggplot(aes(q1, label_int, fill=model, color=model, group=model)) +
#     geom_linerange(aes(x=x, xmin=xmin, xmax=xmax, color=model),
#                    data = . %>% 
#                      group_by(model, label_int) %>% 
#                      summarise(x=mean(q1), xmin=min(q1), xmax=max(q1)),
#                    position = position_dodge(width=.7), color="grey70") +
#     geom_point(aes(shape=targets),
#                position = position_dodge(width=.7)) +
#   scale_fill_manual(values=palettes_model) +
#   scale_color_manual(values = palettes_model_d) +
#   scale_shape_manual(values = c(21,22,23,24)) +
#   labs(y="Node") +
#   guides(color = guide_legend(override.aes = list(pch = 21, size=3)),
#          shape = guide_legend(override.aes = list(size=3))) +
#   theme(legend.position = "bottom", legend.box="vertical") + 
#   theme_bw()
# 
# labeled_sptree <- sptree
# labeled_sptree$tip.label <- tax$Species[match(labeled_sptree$tip.label, tax$mnemo)]
# 
# sptree_plot <- ggtree(labeled_sptree) + 
#   geom_tiplab(aes(label=sub("[^A-Z].+? ", ". ", label)),fontface="italic",
#               align = TRUE) + 
#   geom_nodelab(color="firebrick", 
#                nudge_x = -.15, hjust = 1, 
#                nudge_y = .3, size=3.5)+
#   xlim(c(-4, 21))
# 
# fig_s2 <- (sptree_plot + pp1_plot) +
#   plot_layout(widths = c(1, 1.5)) +
#   plot_annotation(tag_levels = 'A') &
#   theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
# ggsave("draft/figures/figS2.pdf", fig_s2, width = 12, height = 6, device = cairo_pdf)
```

>>>>

p_dbs <- ggplot(jaccard_df, aes(avg_Jaccard, 1, color=singleton, alpha=db)) +
  stat_ecdf(pad=FALSE) +
  scale_color_manual(values=palette_singleton_p) +
  # scale_linetype_manual(values = c(2,2,2,2,1)) +
  scale_alpha_manual(values = seq(0.3, 1, length.out=4)) +
  coord_fixed() +
  scale_y_continuous(expand = expansion(0.0,0)) +  
  scale_x_continuous(expand = expansion(add = c(0,0.01))) +  
  guides(colour = guide_legend(order = 1),  
         alpha = guide_legend(title = "CATH Level", order = 2)) +
  theme(panel.border = element_rect(fill = NA), 
        legend.position = "bottom",
        panel.grid.major = element_line(linetype = 2)) + 
  # theme_bw(base_family = "Aptos") +
  labs(x="Mean Jaccard Index", y="ECDF", 
       alpha="DB", linetype="DB", color="Target sets")

@@

```{r fig3}
xax <- ifelse(df_model$algorithm=="Distance", "italic", "plain")

plot_DL <- reco %>% 
  left_join(df_model) %>% 
  ggplot(aes(weave_factors(model, data), n_events, 
             color=algorithm, fill=targets)) + 
  labs(y="(D+L)/# Tips", x="",fill="Target sets") +
  geom_boxplot(outlier.size = .1) +
  scale_color_manual(values = c("black", "black"),
                     labels = c("ML", expression(italic("Distance")))) +
  guides(x = "axis_nested", 
         colour=guide_legend(title = "Method", override.aes=list(linetype = 0, size=0))) +
  scale_fill_manual(values = palettes_method) +
    theme(
    axis.text.x = element_text(face = xax),
    ggh4x.axis.nestline.x = element_line(color="grey"),
    ggh4x.axis.nesttext.x = element_text(colour = "grey30", 
                                         face = "bold", size=10)
  )

plot_rf <- disco_rf %>% 
  mutate(model=factor(model, levels=models)) %>% 
  filter(nsps>length(sptree$tip.label)/2) %>%
  # separate(bn, c("seed", "targets", "alphabet", "model")) %>% 
  left_join(df_model) %>% 
  ggplot(aes(weave_factors(model, data), rf, 
             color=algorithm, fill=target)) + 
  labs(x="", y="Norm. RF to species tree", fill="Target sets") +
  geom_boxplot(outlier.size = .1) +
  scale_color_manual(values = c("black", "black"),
                     labels = c("ML", expression(italic("Distance")))) +
  guides(x = "axis_nested", 
         colour=guide_legend(title = "Method", override.aes=list(linetype = 0, size=0))) +
  scale_fill_manual(values = palettes_method) +
    theme(
    axis.text.x = element_text(face = xax),
    ggh4x.axis.nestline.x = element_line(color="grey"),
    ggh4x.axis.nesttext.x = element_text(colour = "grey30", 
                                         face = "bold", size=10)
  )

@@

plot_apro <- trees_df %>% 
  left_join(df_model) %>% 
  ggplot(aes(weave_factors(model, data), q1, 
             color=algorithm,fill=targets)) + 
  labs(x="", y="Quartet support", fill="Target sets") + 
  geom_boxplot(outlier.size = .1) +
  scale_color_manual(values = c("black", "black"),
                     labels = c("ML", expression(italic("Distance")))) +
  guides(x = "axis_nested", 
         colour=guide_legend(title = "Method", 
                             override.aes=list(linetype = 0, size=0))) +
  scale_fill_manual(values = palettes_method) +
    theme(
    axis.text.x = element_text(face = xax),
    ggh4x.axis.nestline.x = element_line(color="grey"),
    ggh4x.axis.nesttext.x = element_text(colour = "grey30", 
                                         face = "bold", size=10)
  )

@@

```{r FigS2}
# Compare trees with RF distance
rf_df <- get_rf_df(df_trees)
rf_text <- group_by(rf_df, model.x, model.y) %>% 
  # filter(as.integer(model.x)<=as.integer(model.y)) %>% 
  summarise(value=mean(RF, na.rm=T))

@@

rf_plot <- rf_text %>% 
  ggplot(aes(model.x, model.y)) +
  geom_tile(aes(fill=value)) +
  geom_text(aes(label=round(value, 2)), color="white") +
  scale_fill_distiller(palette="YlGnBu", direction = -1, limits = c(0,1), 
                       name="Mean RF") + 
  labs(x="Mean RF Distance", y="") +
  theme_classic() +
  theme(legend.position = "none", 
        axis.line = element_blank(),
        axis.ticks = element_blank())

@@

fig_s2 <- ((plot_bs / rf_plot) + 
             plot_layout(heights = c(2,1)) | 
             (plot_dist / var_r2t) + 
             plot_layout(heights = c(2,1))) +
  plot_layout(widths = c(1, 2)) +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))

ggsave("draft/figures/Fig_S2.pdf", fig_s2, width = 10, height = 6)
```

# ```{r FigS2}
# pp1_plot <- nodes_df %>%
#     filter(model!="FTPY") %>% 
#     ggplot(aes(q1, label_int, fill=model, color=model, group=model)) +
#     geom_linerange(aes(x=x, xmin=xmin, xmax=xmax, color=model),
#                    data = . %>% 
#                      group_by(model, label_int) %>% 
#                      summarise(x=mean(q1), xmin=min(q1), xmax=max(q1)),
#                    position = position_dodge(width=.7), color="grey70") +
#     geom_point(aes(shape=targets),
#                position = position_dodge(width=.7)) +
#   scale_fill_manual(values=palettes_model) +
#   scale_color_manual(values = palettes_model_d) +
#   scale_shape_manual(values = c(21,22,23,24)) +
#   labs(y="Node") +
#   guides(color = guide_legend(override.aes = list(pch = 21, size=3)),
#          shape = guide_legend(override.aes = list(size=3))) +
#   theme(legend.position = "bottom", legend.box="vertical") + 
#   theme_bw()
# 
# labeled_sptree <- sptree
# labeled_sptree$tip.label <- tax$Species[match(labeled_sptree$tip.label, tax$mnemo)]
# 
# sptree_plot <- ggtree(labeled_sptree) + 
#   geom_tiplab(aes(label=sub("[^A-Z].+? ", ". ", label)),fontface="italic",
#               align = TRUE) + 
#   geom_nodelab(color="firebrick", 
#                nudge_x = -.15, hjust = 1, 
#                nudge_y = .3, size=3.5)+
#   xlim(c(-4, 21))
# 
# fig_s2 <- (sptree_plot + pp1_plot) +
#   plot_layout(widths = c(1, 1.5)) +
#   plot_annotation(tag_levels = 'A') &
#   theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
# ggsave("draft/figures/figS2.pdf", fig_s2, width = 12, height = 6, device = cairo_pdf)
<!-- ``` -->
